---
title: "Phase 2 Dataset QA"
author: "Julia"
date: "2024-06-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Quality Assurance Check

```{r}
# Load the tidyverse (including ggplot2) and janitor
library(tidyverse)
library(janitor)
```

```{r}
# Upload data 
phase2_dat <- read_tsv("Phase2data.txt") |>  janitor::clean_names()
pretest_dat <- read_tsv("pretest_data.txt") |>  janitor::clean_names()
practice_dat <- read_tsv("practice_data.txt") |>  janitor::clean_names()
posttest_dat <- read_tsv("posttest_data.txt") |>  janitor::clean_names()
```

### Goal of this QA:
*Check that all conditions that were supposed to happen happened, that participants completed all conditions (and how many didn't), how many trials per condition per participant, that there is pre-post for all conditions for all participants (and how for how many there isn't), etc.*

```{r}
# check structure of the data
str(phase2_dat) 
```
```{r}
# get summary of the data
summary(phase2_dat) 
```
***It can be seen that the dataset presents 89,110 rows. However, there may be duplicate records that could inflate the participant count. So, a check for duplicates will be done below.***

```{r}
# check for duplicates in the participants
duplicate_check <- phase2_dat |> 
  group_by(anon_student_id, condition) |> 
  summarise(count = n()) |> 
  filter(count > 1)

nrow(duplicate_check)
head(duplicate_check, 12)
```
There are 3,712 duplicates, but if you view the dataset, it can be seen that there are 8 conditions for each duplicated participant. So, there should be around 464 total participants who have completed all 8 conditions.***However, this does not align with our 89,110 total rows in the dataset. This suggests that there might be additional data not accounted for by the 465 participants completing 8 conditons each.***

```{r}
# verify the total number of unique participants after checking duplicates before
total_unique_participants <- phase2_dat |> 
  summarise(unique_participants = n_distinct(anon_student_id))
total_unique_participants
```
***Since there is a mismatch in expected rows and total rows (465 participants, but 89,110 rows), let's find the distribution of entries for participants. The ideal outcome would include most participants with 8 entries. Let's see if there are any anomalies.***

```{r}
# created a dataframe named 'participant_entries' to show the number of unique "entries" or rows for each participant
participant_entries <- phase2_dat |> 
  group_by(anon_student_id) |> 
  summarise(entries = n())

# Summarize the entries to see the distribution
entries_distribution <- participant_entries |> 
  count(entries)

print(participant_entries)
print(entries_distribution)

```
***Overall, it can be seen that there are many participants who have 100+ entries/rows (with the greatest number of participants having 200 entries) which accounts for the high row count of 89,110 in the dataset.***Since there are so many entries per participant, we'll need to count the number of trials per participant, which can be shown a few sections below.

### Check for missing values:

```{r}
# check which columns (variables) have missing values by using 'colSums' function
colSums(is.na(phase2_dat))
```
***From this function, it looks like 'correct_transaction_time', 'step_start_time', 'kc_unique_step', 'opportunity_unique_step', and 'predicted_error_rate_unique_step' have missing values (NAs).***

### Verify all conditions occurred:

```{r}
# outputs all unique conditions present in the 'condition' column to verify that all conditions are present (all conditions that were supposed to happen happened.)
unique(phase2_dat$condition)
```
### Verify if participants completed all conditions: 
```{r}
# check that participants completed all conditions
participant_conditions <- phase2_dat |> 
  group_by(anon_student_id) |> 
  summarise(conditions_completed = n_distinct(condition))

print(paste("Number of participants who completed all conditions:", nrow(participant_conditions))) # count number of participants that completed all conditions

```

```{r}
# check participants who did not complete all conditions
incomplete_participant_cond <- participant_conditions |> 
  filter(conditions_completed < 8)

incomplete_participant_cond
```
***A total of 3 participants did not complete all 8 conditions.***

### Count trials per condition per participant
```{r}
trials_per_condition <- phase2_dat |> 
  group_by(anon_student_id, condition) |> 
  summarise(trials = n())

head(trials_per_condition)
```
***The first row shows that there were 16 trials for the first condition for the first participant. Since they completed all 8 conditions, and each condition has differing numbers of trials, this explains how this first participant has 200 entries/rows, as calculated above.***

### Check Pre-Post Data:
```{r}
# Check for unique participants in each dataset to see if they match up with the Phase 2 dataset
unique_pretest <- n_distinct(pretest_dat$anon_student_id)
unique_posttest <- n_distinct(posttest_dat$anon_student_id)
unique_practice <- n_distinct(practice_dat$anon_student_id)

print(paste("Unique participants in pretest:", unique_pretest))
print(paste("Unique participants in posttest:", unique_posttest))
print(paste("Unique participants in practice:", unique_practice))
```

***It seems like all 464 participants continued throughout the practice, but only 432 participants were able to take part in the post-test.***

```{r}
# merge pre-post-practice datasets
combined_prepost <- bind_rows(pretest_dat, posttest_dat,practice_dat)  

# Check for pre-post data for all conditions for all participants
prepost_check <- combined_prepost |> 
  group_by(anon_student_id, condition) |> 
  summarise(present_test_types = n_distinct(sample)) |>
  filter(present_test_types > 2) 

unique_participants <- prepost_check |> 
  summarise(unique_participants_with_prepost = n_distinct(anon_student_id))

# Check for participants missing pre-post data
missing_prepost <- combined_prepost |> 
  group_by(anon_student_id, condition) |> 
  summarise(present_test_types = n_distinct(sample)) |> 
  filter(present_test_types < 3) |> 
  summarise(unique_missing_prepost = n_distinct(anon_student_id))

print(paste("Number of unique participants with pretest, posttest, or practice data for any condition:", nrow(unique_participants)))
print(paste("Number of unique participants missing pretest, posttest, or practice data for any condition:", nrow(missing_prepost)))
  
```
* Number of unique participants ***with*** pretest, posttest, or practice data for any condition: 431
* Number of unique participants ***missing*** pretest, posttest, or practice data for any condition: 46

```{r}
# check how many trials there were for each condition per student in the pretest, experiment, and posttest
trials_per_condition <- combined_prepost |> 
  group_by(anon_student_id, condition, sample) |> 
  summarise(trials = n(), .groups = 'drop')

head(trials_per_condition) # display only the first 6 rows
```
### Check how many times, across the entire sample of students who completed everything, each question was seen:
```{r}
# first, group by student id and condition to see the distribution of test types and then filter by 'conditions_completed' to only include participants who have completed all conditions and tests 
num_questions <- combined_prepost |> 
  group_by(anon_student_id, condition) |> 
  summarise(present_test_types = n_distinct(sample)) |> 
  filter(present_test_types == 3) |>  # this filters to only show participants who have completed all pre-post tests and practice
  group_by(anon_student_id) |> 
  summarise(conditions_completed = n_distinct(condition)) |> 
  filter(conditions_completed == 8) 

# Get the list of student ids who completed all conditions
students_completed_all <- num_questions$anon_student_id # create a vector containing the participants who have completed all 8 conditions

# Filter the combined data to include only students who completed all conditions
filtered_data <- combined_prepost |> 
  filter(anon_student_id %in% students_completed_all)

# Count the number of times each question was seen
question_counts <- filtered_data |> 
  group_by(problem_name) |>   
  summarise(count = n()) |> 
  arrange(desc(count)) # display count from high to low to see if there is any potential bias

question_counts
```

